Closing the gap between statistical and scientific workflows for improved forecasts in ecology

### 2025-03-28 forecastflows.tex
step back, mentioning machine learning

% Step back
%\item we need more data, and better question (relate this to both previous study cases)
%\item where can we best reduce uncertainties through new scientific insights?
%\item machine learning! If we change nothing, what's the point of not doing ML? ML $>$ process-based without question, and ML $>$ trends without mechanisms! Where theory is lacking, or where we are far from mechanistic understanding, you might as well do ML!
%\item (ML has a way to collect and interpret large datasets...)

>>>>

% Step back
%\item we need more data, and better question (relate this to both previous study cases)
%\item where can we best reduce uncertainties through new scientific insights?
%\item machine learning! If we change nothing, what's the point of not doing ML? ML $>$ process-based without question, and ML $>$ trends without mechanisms! Where theory is lacking, or where we are far from mechanistic understanding, you might as well do ML!
%\item (ML has a way to collect and interpret large datasets...)
Across the different fields of ecology---for both parameter estimation and forecasting---a systematic application of a coherent workflow holds the promise to highlight the opportunities to best reduce uncertainties through new scientific insights, toward the most critical steps. This will help refocus the debate on designing new hypothesis, formulating new questions---and guiding efforts to collect new data. 
In a world where machine learning is rapidly advancing, there is no point of sticking to traditional methods if no changes are made. Machine learning may surpass process-based models if the latter lack a robust estimation of their parameters and fall in a complexity trap, at the cost of their interpretability. Similarly, trend analysis, when the focus is on methodological controversies (due to the lack of an iterative workflow) rather than on a robust mechanistic foundation, offers no clear advantage over machine learning.
% If we do not take care to apply this workflow, machine learning might be the more pragmatic approach...
